Generalized Quasi-linear Approximation and Non-normality in Taylor-Couette Spiral Turbulence

### 2022-08-15 paper/paper.tex

Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda < 30$.
While both $\Lambda 30$ and 60 show spiral turbulence, $\Lambda=30$ has considerably more wave motion in the laminar region than either $\Lambda = 60$ or DNS. 
Interestingly, $\Lambda=30$ has a lower kinetic energy much closer to that of DNS than does $\Lambda=60$.
By examining power spectra (figure
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. $\Reyn_i = 900$ (left), started from small perturbations, is above the linear stability boundary. $\Reyn_i = 700$ (right panel), continued by lower $\Reyn_i$ from previous simulations, is in the hysteretic region below the stability boundary. Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$. At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. At lower $\Reyn_i$, $\Lambda = 60$ is required to maintain the spiral turbulence state. }
    \label{fig:ke_vs_t_rei900}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{../figs/rei900_spectra.pdf}
    \caption{$\theta$-$z$ power spectra of $v$ at midgap for $\Reyn_i = 900$ at various GQL cutoffs. Adding a single GQL mode significantly improves the power spectrum. Spiral turbulence appears at $\Lambda=30$. The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.}
    \label{fig:my_label}
\end{figure}
Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating. As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT). Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches its closest approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.

@@

\subsection{Subcritical spiral turbulence}
In order to study subcritical spiral turbulence, we ran a series of DNS with decreasing $\Reyn_i$. Each is initialised with the last timestep of the prior run and run for one viscous time. Once we reached $\Reyn_i = 700$, we run a DNS for a final viscous time and then repeat the procedure from the supercritical GQL study. The right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$. At $\Lambda = 30$, the solution recovers from the decay and begins to climb in energy again, peaking in a state .

>>>>

Figure~\ref{fig:ke_vs_t_rei900} shows the kinetic energy as a function of time and $\Lambda$ for both the supercritical (left) and subcritical (right) runs. 
The supercritical case shows significant oscillations at all $\Lambda < 30$.
While both $\Lambda = 30$ and 60 show spiral turbulence, $\Lambda=30$ has considerably more wave motion in the laminar region than either $\Lambda = 60$ or DNS. 
Interestingly, figure~\ref{fig:ke_vs_t_rei900} shows that $\Lambda=30$ has a \emph{lower} kinetic energy much closer to that of DNS than does $\Lambda=60$.
By examining power spectra in figure~\ref{fig:rei900_spectra}, we begin to see the reason for the more complex flows seen with $\Lambda > 0$. Even low-order non-linear scatter leads to a steady increase 
\begin{figure}
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei900_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figs/rei700_reo_-3398_KE_vs_t.pdf}
    \end{minipage}
    \caption{Perturbation kinetic energy as a function of time for $\Reyn_o=-3398$. $\Reyn_i = 900$ (left), started from small perturbations, is above the linear stability boundary. $\Reyn_i = 700$ (right panel), continued by lower $\Reyn_i$ from previous simulations, is in the hysteretic region below the stability boundary. Starting from a saturated spiral turbulence state from DNS at $t/\tau_\nu \simeq 1.5$ (left), $t/\tau_\nu \simeq 3$ (right), we continue the solution with QL ($\Lambda = 0$) and GQL with cutoffs $\Lambda = 1,3,5,10,30,60$. At higher $\Reyn_i$, all of the reduced models begin high amplitude oscillations. At lower $\Reyn_i$, $\Lambda = 60$ is required to maintain the spiral turbulence state. }
    \label{fig:ke_vs_t_rei900}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{../figs/rei900_spectra.pdf}
    \caption{$\theta$-$z$ power spectra of $v$ at midgap for $\Reyn_i = 900$ at various GQL cutoffs. Adding a single GQL mode significantly improves the power spectrum. Spiral turbulence appears at $\Lambda=30$. The obvious square in $\Lambda = 60$ represents a sudden drop in spectral power at the cutoff wavenumber.}
    \label{fig:rei900_spectra}
\end{figure}
Focusing on $\Lambda=5$, figure~\ref{fig:rei900_lambda5_story} shows that the flow goes through a series of different states bearing a strong resemblance to other well-known TC flow patterns originally identified in exhaustive series of experiments reported in \cite{1984JFM...146...45M} and subsequently found in simulations (e.g. \cite{2009PhRvE..80d6315M}). At the time of kinetic energy minimum, labelled 1 in the left hand side of figure~\ref{fig:rei900_lambda5_story}, the flow seems to recapitulate the interpenetrating spirals (IPS) characteristic of lower $\Reyn$ counter-rotating. As the kinetic energy rises, reaching a knee at point 2, the flow morphology resembles the patchy bursts characteristic of the intermittency regime (INT). Finally, at the peak of the kinetic energy oscillation, point 3 in figure~\ref{fig:rei900_lambda5_story}, the flow reaches its closest approach to something resembling spiral turbulence, though it is very difficult to identify the flow as such from a slice at a single point in time and plane in space.

@@

\subsection{Subcritical spiral turbulence}
In order to study subcritical spiral turbulence, we ran a series of DNS with decreasing $\Reyn_i$. Each is initialised with the last timestep of the prior run and run for one viscous time. Once we reached $\Reyn_i = 700$, we run a DNS for a final viscous time and then repeat the procedure from the supercritical GQL study. The right panel of figure~\ref{fig:ke_vs_t_rei900} shows  that all $\Lambda < 30$ revert to the laminar state within $0.1 \tau_\nu$. At $\Lambda = 30$, the solution recovers from the decay and begins to climb in energy again but ultimately returns to the laminar CCF state. Curiously, our $\Lambda = 60$ run not only remains in a non-laminar solution for as long as we have been able to run for, it also correctly reproduces the spiral turbulence pattern, just as in the supercritical case.

The ability of GQL to reproduce spiral turbulence in the subcritical regime is quite significant. First, while $\Lambda = 60$ corresponds to the fully non-linear interactions occurring in a rather large fraction of the total number of modes in the simulation, it is important to note that a DNS at $n_\theta = n_z = 60$ would be massively underresolved, possibly numerically unstable. Second, the ability of non-linear interactions to sustain turbulence where non-normal linear dynamics alone cannot suggests an important role for exact coherent states in subcritical Taylor-Couette flow. Because GQL retains all linear dynamics and low-order non-linear couplings, in addition to the couplings between the mean flows and high modes, if non-normal dynamics played a strong role in subcritical turbulence, one would expect the behavior seen at $\Lambda = 30$ to lead to sustained, rather than transient turbulence.
